### 2016-02-05 Manuscript/BiologyLetters/response.tex

\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\setlength\parindent{0pt}

\begin{document}

\textbf{RE: Decision on Manuscript ID RSBL-2015-1003}\\
\bigskip
Dear Surayya Johar,\\
\bigskip
We are very grateful to the four referees for their helpful and constructive comments.s
We have taken all of their comments on board, and respond to their points below. To improve the clarity of this document, we dealt with each comment in the order they appeared in your ``Decision on Manuscript'' email and we have included the reviewers comments in blue.

@@

We replaced the Table 1 in the main text with the Table 1 in the ESM 2 (available data without any number of characters threshold - see reviewer 2 comment 1).
We therefore removed ESM 2 and renamed ESM 3 to ESM 2

\end{enumerate}

%%%%%%%%%%%%%%%
% Reviewer 2 (Graham Slater)
%%%%%%%%%%%%%%%

\section{Reviewer 2 (Graham Slater):}
\begin{enumerate}
\item{\textcolor{blue}{The authors state on page 3 lines 82-90 that they avoided matrices with ``few characters'' due to lower probability of overlapping characters with other matrices and so only selected matrices containing >100 characters. First, a lack of overlapping characters doesn’t seem justifiable, at least to my mind, as a reason to exclude a matrix if the ultimate goal is to infer the placement of extinct taxa relative to extant taxa. Antler ramifications, to use the authors' example, are likely to be a very useful character when attempting to resolve relationships among living and extinct artiodactyls, even if they cannot be coded for monotremes. To flip the argument, a matrix of 200 + characters for mysticetes is similarly likely to contain many non-applicable characters from the perspective of talpids. I do understand that the authors are thinking downstream in terms of super matrices for higher level analyses, but smaller matrices are going to be just as relevant for reaching this goal. You don't have much space in the ms I understand, but surely this needs to be addressed?!}}

%TO DO: let's just remove this threshold thingy since I seem to be the only one to be convinced by it ;)...
% NC: Yep that seems sensible
We removed the analysis with a threshold of OTUs with at least 100 characters and replaced it with the analysis without any minimal number of morphological characters.
We retracted the paragraph aout the character threshold analysis and updated the results of figure 1 and table 1 to contain all living OTUs; we removed the redundant ESM 2 and we updated the figures

@@

\item{\textcolor{blue}{Line 104 - ``NTI 15'' should probably be ``NTI [15]''.}}
%TO DO IN THE WORD MANUSCRIPT! 
We fixed this typo.

\item{\textcolor{blue}{Line 104 - It's not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
We have clarified this by changing ``observed mean distance between each of $n$ taxa'' to ``observed mean sum of the branch lengths between each of $n$ taxa''. % NC: Fix in document %TG:done

\item{\textcolor{blue}{Line 117 - It's not clear to me how supraspecific values were calculated if the reference phylogeny (ref 10) is species-level. Are you using the node subtending the higher taxon or something else?}}
%TO CHECK!
% NC: Probably you've pruned the phylogeny to include only species within each order/family?

\item{\textcolor{blue}{Line 149 - ``most palaeontological studies use the genus as their smallest OTU'' I'm not sure if this is strictly true. It's more that shorter OTU names are commonly used. After all at least one actual specimen has to be coded and in my experience if you read the paper usually only a single species is used. Palaeontologists are just bad at using proper binomials in their matrices (which is an issue, but a different one). In addition ref 20 is actually all species-level OTUs! (I didn't check ref 21.)}}
%To CHECK, do we actually care? 
% NC: Just change it from "most" to "many", or say paleo studies commonly feature only one species in a genus.
We modified this statement by stating that ``many palaeontological studies use the genus (or monospecific genera) as their smallest OTU''.

### 2016-02-05 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Updated results in text (no threshold)

%---------------------------------------------
%
%       METHODS
%
%---------------------------------------------
\section{Materials and Methods}
\subsection{Data collection and standardisation}
We downloaded all cladistic matrices %TG: I left "cladistic" here 'cause that's what the matrix is (even folloowing Peter's definition)!
 containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
We used OTUs rather than species because entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.

Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not applicable% rev 1.2.4
 across large clades (e.g. ``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
Results of analyses with no threshold are available in the ESM. 
After removing matrices with \textless 100 characters, we retained 1074 unique living OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)

@@

\section{Results}
22 of 28 orders % Change that!
 have low coverage (\textless 25\% species with available anatomical characters) and six have high coverage %Change that!
  (\textgreater 75\% species with available anatomical characters) at the species-level.
At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table\ref{Table_results}).

@@

Only six orders had significantly clustered data (Afrosoricida and Pholidota at the species-level, and Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).

Figure \ref{Figure_example_coverage} shows randomly distributed OTUs with cladistic data in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae; Figure \ref{Figure_example_coverage}B).

@@

The number of living mammalian OTUs with no available anatomical characters was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available anatomical characters.
This high coverage threshold of 75\% of taxa with available anatomical characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
Therefore we expect difficulties in placement of fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
This point is important from a practical point of view because of the slight discrepancy between neontological and palaeontological species concepts.
While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
Therefore, many palaeontological studies use the genus (or monospecific genera) as their smallest OTU (e.g. \cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.

When few species have available anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil branching in the right clade.
The second best scenario is that species with available anatomical characters are randomly distributed across the phylogenies. 
Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available anatomical characters were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with available anatomical characters are phylogenetically clustered. 
Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
, and will have more chance to randomly branch within Canidae (Figure \ref{Figure_example_coverage}B).

>>>>

%---------------------------------------------
%
%       METHODS
%
%---------------------------------------------
\section{Materials and Methods}
\subsection{Data collection and standardisation}
We downloaded all cladistic matrices %TG: I left "cladistic" here 'cause that's what the matrix is (even following Peter's definition)!
 containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
We used OTUs rather than species because entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.

% Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not applicable% rev 1.2.4
%  across large clades (e.g. ``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
% Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
% This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
% Results of analyses with no threshold are available in the ESM. 
% After removing matrices with \textless 100 characters, we retained 1074 unique living OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)
We retained 1601 unique living OTUs from 286 matrices.

@@

\section{Results}
11 out of 28 orders have low coverage (\textless 25\% species with available anatomical characters) and 7 have high coverage (\textgreater 75\% species with available anatomical characters) at the species-level.
At the genus-level, one order have low coverage and 15 have high coverage, and at the family-level, no orders have low coverage and 25 have high coverage (Table\ref{Table_results}).

%TG: former results:
% 22 of 28 orders have low coverage (\textless 25\% species with available anatomical characters) and six have high coverage (\textgreater 75\% species with available anatomical characters) at the species-level.
% At the genus-level, three orders have low coverage and 12 have high coverage, and at the family-level, no orders have low coverage and 23 have high coverage (Table\ref{Table_results}).

@@

Only seven orders had significantly clustered data (Cetartiodactyla, Cingulata, Pilosa and Rodentia at the species-level and Carnivora, Chiroptera and Soricomorpha at both species- and genus-level) and none had significantly overdispersed data (Table~\ref{Table_results}).

Figure \ref{Figure_example_coverage} shows randomly distributed OTUs with cladistic data in Primates (Figure \ref{Figure_example_coverage}A) and phylogenetically clustered OTUs with cladistic data in Carnivora (mainly Canidae and Urisdae but no Herpestidae; Figure \ref{Figure_example_coverage}B).

@@

The number of living mammalian OTUs with no available anatomical characters was surprisingly high at the species-level: only seven out of 28 orders have a high coverage of taxa with available anatomical characters.
This high coverage threshold of 75\% of taxa with available anatomical characters represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
Therefore we expect difficulties in placement of fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
This point is important from a practical point of view because of the slight discrepancy between neontological and palaeontological species concepts.
While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
Therefore, many palaeontological studies use the genus (or monospecific genera) as their smallest OTU (e.g. \cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.

When few species have available anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil branching in the right clade.
The second best scenario is that species with available anatomical characters are randomly distributed across the phylogenies. 
Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available anatomical characters were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with available anatomical characters are phylogenetically clustered. 
Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
, and will have more chance to randomly branch within Canidae or Ursidae (Figure \ref{Figure_example_coverage}B).
